---
title: "RepResearch"
author: "Kaushank"
date: "Thursday, February 12, 2015"
output:
  html_document:
    keep_md: yes
  pdf_document:
    keep_tex: yes
---
Reproducible Research: Peer Assessment 2

<b>1. Assignment</b>

This assignment explores the  NOAA Storm Database and attempts to answer some basic questions about severe weather events. 


<b>2. Synopsis></b>
NOAA maintains a public database for storm events depicting details like location,  estimates for damage to property,damage to life and property etc..
In this report I am attempting to  investigate which events have greater impact 
:Financial,Fatalities,Injuries, Damage to Property and Damage to Crops.

Conclusion:
From the analysis, most  physical phenomena result in injuries to people, which sometimes are fatal. By far, Tornadoes are the most dangerous events, caused ~100000 injuries on the last 60 years.

When analysing the event types by the impact on the economy, we observe that floods caused $15 billions damages on the last 60 years, mostly on properties.

Tornado's are by far the highest cause for injuries, and second in fatalities, whilst heat &amp; drought cause the most fatalities, but fourth in injuries. Both are in the top 5 of injuries &amp; fatalities next to Thunderstorms. Flooding  and Snow &amp; Ice . In economic damages, only the property damage really factors in the total damage, except for Heat &amp; Drought where more than 90% of damages is determined by crop damage. The #1 &amp; #2 of weather damage sources, resp. Flooding &amp; High Surf and Wind &amp; Storm cover more than 80% of all economic cost, while Wind &amp; Storm aren't even in the top 5 of victims.</p>


<b>3.Data Processing</b>

<b>3.1 Set the appropriate Working Directory </b>
Change the working Directory to Project Directory for Assignment 2
```{r}
setwd("C:/Users/Dell1/Desktop/DSSGProjects/RR/RepData_PeerAssessment2")
```

<b>3.2 Load the necessary Librarries</b>

```{r, echo=FALSE}
library(R.utils) # for bunzip2 algorithm compression
library(plyr) # for count & aggregate method
library(reshape2) # for melt 
library(ggplot2) # for plots
library(scales) # for plot scaling
library(Hmisc)
library(knitr)
library(markdown)
library(car)

```

<b>3.2. Load source file and extract it</b>
Unzip the data

```{r}
stormdata <- read.csv("./StormData.csv")
damages<-aggregate(cbind(FATALITIES, INJURIES) ~ EVTYPE , stormdata, sum)
dam<-melt(head(damages[order(-damages$FATALITIES,-damages$INJURIES),],10))

stormdata$PROPDMG<-stormdata$PROPDMG*as.numeric(Recode(stormdata$PROPDMGEXP, "'0'=1;'1'=10;'2'=100;'3'=1000;'4'=10000;'5'=100000;'6'=1000000;'7'=10000000;'8'=100000000;'B'=1000000000;'h'=100;'H'=100;'K'=1000;'m'=1000000;'M'=1000000;'-'=0;'?'=0;'+'=0",as.factor.result=FALSE))
stormdata$CROPDMG<-stormdata$CROPDMG*as.numeric(Recode(stormdata$CROPDMGEXP, "'0'=1;'2'=100;'B'=1000000000;'k'=1000;'K'=1000;'m'=1000000;'M'=1000000;''=0;'?'=0",as.factor.result=FALSE))

economic<-aggregate(cbind(PROPDMG, CROPDMG) ~ EVTYPE , stormdata, sum)
eco<-melt(head(economic[order(-economic$PROPDMG,-economic$CROPDMG),],10))

ggplot(dam, aes(x=EVTYPE,y=value,fill=variable)) + geom_bar(stat = "identity") + coord_flip() +
  ggtitle("Harmful events") + labs(x = "", y="number of people impacted") +
  scale_fill_manual (values=c("blue","red"), labels=c("Deaths","Injuries"))

ggplot(eco, aes(x=EVTYPE,y=value,fill=variable)) + geom_bar(stat = "identity") + coord_flip() +
  ggtitle("Economic consequences") + labs(x = "", y="cost of damages in dollars") +
  scale_fill_manual (values=c("yellow","green"), labels=c("Property Damage","Crop Damage"))

```
